---
title: "README"
author: "Eric Graves"
date: "October 12, 2016"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(isofor)
```

## Isolation Forest

An Isolation Forest is an ensemble of completely random decision trees. At each split a random
feature and a random split point is chosen. Anomalies are isolated if they end up in a partition
far away from the rest of the data. In decision tree terms, this corresponds to a record that 
has a short "path length". The path length is the number of nodes that a record passes through
before terminating in a leaf node. Records with short average path lengths through the entire
ensemble are considered anomalies.

## An analogy

Describing the location of a country home takes many fewer directions than describing the 
location of a brownstone in Brooklyn. The country home might be described as "the only house
on the south shore of Lake Woebegon". While directions to the brownstone must be qualified
with much more detail: "Go north on 5th Street for 12 blocks, take a left on Van Buren, etc.."

The country house in this example is a literal outlier. It is off by itself away from most other
homes. Similarly, records that can be described succinctly are also outliers.

## Example

Here we create two random, normal vectors and add some outliers.

```{r dummy-data}

N = 1e3
x = c(rnorm(N, 0, 0.5), rnorm(N*0.05, -1.5, 1))
y = c(rnorm(N, 0, 0.5), rnorm(N*0.05,  1.5, 1))
data = data.frame(x, y)
plot(data)
title("Dummy data with outliers")
```



```{r isofor}
mod = iForest(X = data, 100, 32)
p = predict(mod, data)
col = ifelse(p > quantile(p, 0.95), "red", "blue")
plot(x, y, col=col)
```

```{r kmeans}
km = kmeans(data, 2)
plot(x, y, col=km$cluster+1)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
